---
title: "Concepts of ggplot2"
output: html_notebook
---

# Loading packages

```{r}
packages <- c(
    "tidyverse",
    "ggplot2"
)

package_check <- lapply(
  packages,
  function(x) {
    if (!require(x, character.only = TRUE)) {
      install.packages(x, dependencies = TRUE)
      library(x, character.only = TRUE)
    }
  }
)
```

# The grammer of ggplot2
```{r}
tibble(
Component = c("1-Data","2-Aesthetics", "3-Geometries", "4-Statistics", 
              "5-Scales", "6-Coordinate System", "7-Facets", "8-Theames"), 
Function =  c("ggplot(data)", "aes()", "gemo_*()", "stat_*()", "scale_*()", 
              "coord_*()", "facet_*()", "theame_*()"),
Description = c("The raw data you want to visualize",
                "Aesthetic mappings between variables and visual properties", 
                "The geometric shapes representing the data",
                "The statistics transformation applied to the data",
                "Maps between the data and the asesthetic dimmentions",
                "Maps data into the plane of the data rectengle",
                "The arrangement of the data into a grid of plots",
                "The overall visiual of the garaph"
                )
)
```
# Load Data


```{r}
bikes_df <- read_csv(file.path("data", "london-bikes-custom.csv"),
  col_types = "Dcfffilllddddc"
)
```
## 1- Data 
initiate an empty canvas
```{r, echo=FALSE, fig.width=4, fig.height=3}
ggplot(data = bikes_df)
```
## 2- Aesthetics
Maps data (variables) to the graph properties such as;

* positions(x,y)
* colours(colours, fill)
* shapes(shape, linetype)
* size(size)
* transparent(alpha)
* groupings(group)

```{r, echo=FALSE, fig.width=4, fig.height=3}
bikes_df %>%
  ggplot(aes(x = temp_feel, y = count))
```
## 3- Geometries
Interpret data (aesthetics) as graphs.

* points
* lines
* polygons
* text labels
* ...

```{r, echo=FALSE, fig.width=4, fig.height=3}
bikes_df %>%
  ggplot(aes(x = temp_feel, y = count)) +
  geom_point()
```
we may add visual properties to the graph
```{r, echo=FALSE, fig.width=4, fig.height=3}
bikes_df %>%
  ggplot(aes(x = temp_feel, y = count)) +
  geom_point(
    color = "#2F005F",
    alpha = .5,
    shape = "X",
    stroke = 1,
    size = 2
  )
```
### Mapping vs visual properties

Mapping 

```{r, echo=FALSE, fig.width=4, fig.height=3}
ggplot(
    bikes_df,
    aes(x = temp_feel, y = count)
  ) +
  geom_point(
    color = "#28a87d",
    alpha = .5
  )
```
 Visual properties
 
```{r, echo=FALSE, fig.width=4, fig.height=3}
ggplot(
    bikes_df,
    aes(x = temp_feel, y = count)
  ) +
  geom_point(
    aes(color = season),
    alpha = .5
  )

```
### Mapping expresions

```{r}
ggplot(
    bikes_df,
    aes(x = temp_feel, y = count)
  ) +
  geom_point(
    aes(color = temp_feel > 20),
    alpha = .5
  )
```
### Global vs Local

```{r}
ggplot(
    bikes_df,
    aes(x = temp_feel, y = count,
        color = season)
  ) +
  geom_point(
    alpha = .5
  )
```
### DIV
<div class='fluid-row'>
  <div class='col-md-6'>

1st column when browser is wide

```{r}
ggplot(
    bikes_df,
    aes(x = temp_feel, y = count,
        color = season)
  ) +
  geom_point(
    alpha = .5
  )
```
  </div>
  <div class='col-md-6'>

... 2nd column when browser is wide

```{r}
ggplot(
    bikes_df,
    aes(x = temp_feel, y = count)
  ) +
  geom_point(
    aes(color = temp_feel > 20),
    alpha = .5
  )
```
  </div>

</div>
# DIV format
<div class='fluid-row'>
<div class='col-md-6'>
  1st column code
```{r}
ggplot(
    bikes_df,
    aes(x = temp_feel, y = count,
        color = season)
  ) +
  geom_point(
    alpha = .5
  )
```
  
</div>
<div class='col-md-6'>
  2st column code
```{r}
ggplot(
    bikes_df,
    aes(x = temp_feel, y = count,
        color = season)
  ) +
  geom_point(
    alpha = .5
  )
```
  
</div>
</div>

<div class='fluid-row'>
<div class='col-md-6'>
  1st column plot
```{r echo=FALSE}
ggplot(
    bikes_df,
    aes(x = temp_feel, y = count,
        color = season)
  ) +
  geom_point(
    alpha = .5
  )
```

</div>
<div class='col-md-6'>
  2st column plot
```{r echo=FALSE}
ggplot(
    bikes_df,
    aes(x = temp_feel, y = count,
        color = season)
  ) +
  geom_point(
    alpha = .5
  )
```
  
</div>
</div>
# END

```{r echo=FALSE}
ggplot(
    bikes_df,
    aes(x = temp_feel, y = count,
        color = season)
  ) +
  geom_point(
    alpha = .5
  )
```

```{r echo=FALSE}
ggplot(
    bikes_df,
    aes(x = temp_feel, y = count)
  ) +
  geom_point(
    alpha = .5
  )
```

